---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Keng (庚)

<!-- badges: start -->
<!-- badges: end -->

The `Keng` package is named after Loo-Keng Hua, who made great achievements in mathematics mainly through self-study. Hua's spirit encourages Qingyao, and inspires the `Keng` package.

The `Keng` package aims to automate the computations Qingyao repeat in his research and teaching. Hope the functions and data gathered in this package help to ease your life.

## Installation

You can install the development version of `Keng` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pkg_install("qyaozh/Keng")
```

## Load

Before using the `Keng` package, load `Keng` using `library()` function.

```{r library}
library(Keng)
```

## List of contents

Here is a list of the functions and data gathered in the `Keng` package. Their Useges are detailed in the documentation.

### Pearson's *r*

`cut_r()` give you the cut-off values of Pearson's *r* at the significance levels of *p* = 0.1, 0.05, 0.01, 0.001 when the sample size *n* is known.

`test_r()` tests the significance of *r* using *t*-test when *r* and *n* is known.

### PRE of `lm()`

`test_PRE()` computes *PRE* of the augmented model compared with the compact model and test it using the *F* test.
