---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Keng

<!-- badges: start -->
<!-- badges: end -->

The goal of `Keng` is to automate the computations Qingyao repeat in his research and teaching. Hope these functions ease your life too.

## Installation

You can install the development version of `Keng` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("qyaozh/Keng")
```

## Example

Before using these functions `Keng` gathered, load `Keng` using `library()` function.

```{r library}
library(Keng)
```

`cut_r()` give you the cut-off values of Pearson's *r* at the significance levels of *p* = 0.1, 0.05, 0.01, 0.001 when the sample size *n* is known. 

```{r cut_r}
cut_r(193)
```

`r` with the absolute value larger than the cut-off value is significant at the corresponding significance level. As the results above suggests, when *n* = 193, |r| > 0.1412906 is significant at the significance level of 0.05 (e.g., ±0.15).

```{r test_r}
test_r(0.2, 193)
```

`test_r()`  test the significance of *r* using *t*-test when *r* and *n* is known.
